---
title: "Class4"
author: "Emily"
date: "February 13, 2018"
output: html_document
language: r 
cache: packages 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<bold>Introduction:</bold> 

This is a project to examine speech data from a study that followed the symptom progression of patients with Parkinsons disease. 

<bold>Data:</bold> 

Original paper: A Tsanas, MA Little, PE McSharry, LO Ramig (2009)
'Accurate telemonitoring of Parkinson.s disease progression by non-invasive 
speech tests',
IEEE Transactions on Biomedical Engineering (to appear). 

Retrieved from UCL Machine Learning Database: https://archive.ics.uci.edu/ml/datasets/Parkinsons+Telemonitoring

ATTRIBUTE INFORMATION:

subject# - Integer that uniquely identifies each subject
age - Subject age
sex - Subject gender '0' - male, '1' - female
test_time - Time since recruitment into the trial. The integer part is the 
number of days since recruitment.
motor_UPDRS - Clinician's motor UPDRS score, linearly interpolated
total_UPDRS - Clinician's total UPDRS score, linearly interpolated
Jitter(%),Jitter(Abs),Jitter:RAP,Jitter:PPQ5,Jitter:DDP - Several measures of 
variation in fundamental frequency
Shimmer,Shimmer(dB),Shimmer:APQ3,Shimmer:APQ5,Shimmer:APQ11,Shimmer:DDA - 
Several measures of variation in amplitude
NHR,HNR - Two measures of ratio of noise to tonal components in the voice
RPDE - A nonlinear dynamical complexity measure
DFA - Signal fractal scaling exponent
PPE - A nonlinear measure of fundamental frequency variation 

```{r}
library(DataPrep)
data("parkinsons_updrs_data") #retrieves from data folder

library(DT)
datatable(head(parkinsons_updrs_data))
```

<bold>Research Question:</bold> 

Is there a correlation between increased jitter and increased shimmer in speech data among Parkinson's patients? 

<bold>Reseach Plan:</bold> 

1. Run a linear model of shimmer and jitter
```{r Linear Model}
myModel = lm(parkinsons_updrs_data$`Jitter(Abs)`~ parkinsons_updrs_data$Shimmer)

```

Session Info: 
```{r}
 session_info()
```

